---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data

Let's load data from zip file using readr library.
```{r echo=TRUE}
library("readr")
dataset <- read_delim("activity.zip",delim=",")

print(dataset)
```



## What is mean total number of steps taken per day?
We are going to group by date and do a sum of steps per date.

```{r echo=TRUE}
library("dplyr")

sum_by_date <- dataset %>% group_by(date) %>% summarize_all(sum)

```


Then we use ggplot2 to plot number of steps each day.
```{r echo=TRUE}
library("ggplot2")

p <- ggplot(sum_by_date, aes(x=date,y=steps) ) 
p <- p + geom_col()
p <- p + xlab("Dates") + ylab("Steps") 
p <- p + ggtitle("Sum of steps per day") 

print(p)
```

Now let's find mean and median
```{r echo=TRUE}
my_mean <- mean(sum_by_date$steps,na.rm=TRUE)
my_median <- median(sum_by_date$steps,na.rm=TRUE)
print(my_mean)
print(my_median)

```

Mean for total steps per day is `r format(my_mean,nsmall=2)`.  
Median for total steps per day is `r format(my_median,nsmall=2)`.  



## What is the average daily activity pattern?



## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
