---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
### Let's load libraries here

```{r echo=TRUE}
library("readr")
library("dplyr")
library("ggplot2")
library("gridExtra")

```

### Data load

Let's load data from zip file using readr library.
```{r echo=TRUE}
dataset <- read_delim("activity.zip",delim=",")

print(dataset)
```



## What is mean total number of steps taken per day?

### Initial grouping
We are going to group by date and do a sum of steps per date.

```{r echo=TRUE}

sum_by_date <- dataset %>% group_by(date) %>% summarise(sum_steps=sum(steps, na.rm=TRUE))

```

### Plot sum of steps per day
Then we use ggplot2 to plot number of steps each day.
```{r echo=TRUE}

p <- ggplot(sum_by_date, aes(x=date,y=sum_steps) ) 
p <- p + geom_col()
p <- p + xlab("Dates") + ylab("Steps") 
p <- p + ggtitle("Sum of steps per day") 

print(p)
```

### Mean and Median of sum of steps per day
Now let's find mean and median
```{r echo=TRUE}
my_mean <- mean(sum_by_date$sum_steps,na.rm=TRUE)
my_median <- median(sum_by_date$sum_steps,na.rm=TRUE)
print(my_mean)
print(my_median)

```

Mean for total steps per day is `r format(my_mean,nsmall=2)`.  
Median for total steps per day is `r format(my_median,nsmall=2)`.  



## What is the average daily activity pattern?

### Grouping by interval
We are going to group by interval and find the average number of steps for each interval.
As there are 288 intervals of 5 minutes in a day (60*24 minutes / 5 = 288).
We should find 288 interval in the generated data.

```{r echo=TRUE}

avg_by_interval <- dataset %>% group_by(interval) %>% summarise(avg_steps=mean(steps, na.rm=TRUE))
summary(avg_by_interval)
print(avg_by_interval)

```


### Plot sum steps by interval
Let's plot this data in a time series plot

```{r echo=TRUE}

p <- ggplot(avg_by_interval, aes(x=interval,y=avg_steps) ) 
p <- p + geom_col()
p <- p + xlab("Intervals") + ylab("Steps") 
p <- p + ggtitle("Mean of steps per interval") 

print(p)

```


### Maximum
Let's find the maximum average steps.

```{r echo=TRUE}

max_steps <- max(avg_by_interval$avg_steps)
print(max_steps)

max_row <- which.max(avg_by_interval$avg_steps)
max_interval <- as.integer(avg_by_interval[max_row,][,1])


```

The maximum avergae steps per interval is  `r format(max_steps,nsmall=2)`.  
The interval when it occures is `r format(max_interval)`. 


## Imputing missing values

### Is there any missing values ? 
The dataset containes missing values:  

```{r echo=TRUE}
missing <- is.na(dataset$steps)

n_missing <- sum(missing)

```

The total missing values for steps is `r format(n_missing)`.

### Let's fill the gapes !

We can replace these missing values by mean of the same interval. We allready have theses values from avg_by_interval.   
Let's create a new dataset, with a new column named new_steps.  

```{r echo=TRUE}

newdataset <- dataset %>% mutate(new_steps=ifelse(is.na(steps),pull(avg_by_interval[interval,2]),steps) )

print(newdataset)

```


### Sum of steps per day 

Let's create a new group by dataset.
```{r echo=TRUE}

sum_by_date <- newdataset %>% group_by(date) %>% summarise(sum_steps=sum(new_steps, na.rm=TRUE))
print(sum_by_date)

```


Then we use ggplot2 to plot number of steps each day.
```{r echo=TRUE}

p <- ggplot(sum_by_date, aes(x=date,y=sum_steps) ) 
p <- p + geom_col()
p <- p + xlab("Dates") + ylab("Steps") 
p <- p + ggtitle("Sum of steps per day") 

print(p)
```


### impact on Mean and Median 
Now let's find mean and median
```{r echo=TRUE}
my_new_mean <- mean(sum_by_date$sum_steps,na.rm=TRUE)
my_new_median <- median(sum_by_date$sum_steps,na.rm=TRUE)
print(my_new_mean)
print(my_new_median)

```

Mean for total steps per day is now `r format(my_new_mean,nsmall=2)` as it was `r format(my_mean,nsmall=2)`.  
Median for total steps per day is now `r format(my_new_median,nsmall=2)` as it was `r format(my_median,nsmall=2)`.  


The mean is changed by the following percentage: `r format(   100*(my_new_mean-my_mean)/my_mean   ,digits=3,nsmall=2)`.
The impact is small.










## Are there differences in activity patterns between weekdays and weekends?

### Create a new factor column containing 'day type'
Let's create this new column with weekday or weekend values.

```{r echo=TRUE}
week_days <- c('Monday','Tuesday','Wednesday','Thursday','Friday')
weekend_days <- c('Saturday','Sunday')

new <- newdataset %>% mutate(day = factor( case_when( weekdays(date) %in% week_days ~ "Weekday",  weekdays(date) %in% weekend_days ~ "Weekend") )) 

```

And convert it to a factor.

```{r echo=TRUE}
new$day <- as.factor(new$day)
print(new)

```

Let's group steps per day and interval and apply the mean to this groups.

```{r echo=TRUE}
avg_by_interval <- new %>% group_by(interval,day) %>% summarise(avg_steps=mean(new_steps, na.rm=TRUE))
summary(avg_by_interval)
print(avg_by_interval)

```


### Let's plot all that

We will use ggplot2 with 2 facets, using day factor.

```{r echo=TRUE}
g <- ggplot(avg_by_interval, aes(x=interval,y=avg_steps) ) 
g <- g + geom_line(alpha=1/2, color="blue") 
g <- g + facet_wrap( . ~ day, nrow=2, ncol=1 ) 
g <- g + theme_gray(base_family = "Avenir", base_size=10)
g <- g + labs(x=" Interval ") + labs(y="Number of Steps" ) 

print(g)
```



